---
title: "EV Power - Lab 4 Project Report Anjali"
format: typst
---

# Example Solution 1


## **Part 0: libraries**

```{r}
install.packages("tidyverse")
library(tidyverse)
library(readr)
library(ggplot2) 
```

## **Part 1:** **Defining Research Question**

Chosen Question: Which states have the highest share of renewable energy? Does that correlate with EV registrations? 

## **Part 2: Data Preparation and Cleaning**

```{r}
clean_names <- function(df) {
  df |>
    rename_with(~ .x |> 
    str_to_lower() |>
    str_replace_all("[^a-z0-9]+", "_") |>
    str_replace_all("^_|_$", ""))
}

state_lookup <- tibble(
  state_full = state.name,
  abb = state.abb
) |> 
    add_row(state_full = "District Of Columbia", abb = "DC") |>
    add_row(state_full = "United States", abb = "US")

ev_use <- read.csv("data/ev-registrations-by-state-2023.csv") |>
  clean_names() |>
  slice(-(1:2)) |> 
  rename(
    state = electric_vehicle_registrations_by_state_2023,
    ev_count_2023 = x 
  ) |>
    mutate(ev_count_2023 = str_extract(ev_count_2023, "\\d+\\.?\\d*") |> 
    as.numeric()) 
head(ev_use)

renew_use <- read.csv("data/renew-use-2023.csv") |>
  clean_names() |>
  mutate(state_raw = state,
  state_abbrev = if_else(str_detect(state_raw, "^[A-Za-z]{2}$"), 
  str_to_upper(state_raw),
  NA_character_)
  ) |>
  left_join(state_lookup, by = c("state_abbrev" = "abb")) |>
  mutate(state = coalesce(state_full, str_to_title(state_raw)), renewable_use_2023 = readr::parse_number(renewable_use_2023)
  ) |>
  select(state, energy_source, renewable_use_2023)
renew_use_summary <- renew_use |>
  group_by(state) |>
  summarize(total_renewable_2023 = sum(renewable_use_2023, na.rm = TRUE), .groups = "drop")

head(renew_use)

```

## **Part 3: Joining / Pivoting Datasets for Analysis**

```{r}
joined <- renew_use_summary |>
  left_join(ev_use, by = "state") |>
  filter(state != "United States")

top_renew <- joined |>
  arrange(desc(total_renewable_2023)) |>
  slice_head(n = 10)

cat("Top 10 States by Renewable Energy (2023):\n")
print(top_renew[, c("state", "total_renewable_2023")])
summary(joined)
head(joined)
```

## **Part 4: Mapping Visualization**

The datasets used include:
- State renewable energy generation (kWh)
- EV registration data (2023)
After cleaning and joining these datasets, we computed total renewable energy per state and compared it to EV counts.

```{r}
library(ggplot2)

ggplot(joined, aes(x = total_renewable_2023, y = ev_count_2023)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Renewable Energy Use vs EV Registrations (2023)",
    x = "Total Renewable Energy (kwh)",
    y = "Number of Registered EVs"
  ) +
  theme_minimal()

ggplot(top_renew, aes(x = reorder(state, total_renewable_2023), y = total_renewable_2023)) +
  geom_col(fill = "dark green") +
  coord_flip() +
  labs(
    title = "Top 10 States by Renewable Energy Generation (2023)",
    x = "State",
    y = "Total Renewable Energy"
  ) +
  theme_minimal()
```

## **Part 5: Interpretation**
States with higher renewable generation (e.g., California, Texas) also tend to have more EVs, suggesting a positive relationship. However, this correlation is not perfect — some states with high renewable capacity have fewer EVs, possibly due to differences in policy or infrastructure.

